---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- badges: start -->

# [geocoder](https://layik.github.io/geocoder/) &middot; [![Travis build
status](https://travis-ci.org/layik/geocoder.svg?branch=master)](https://travis-ci.org/layik/geocoder) [![codecov](https://codecov.io/gh/layik/geocoder/branch/master/graph/badge.svg)](https://codecov.io/gh/layik/geocoder)

Use the power of MongoDB to make spatial queries, find boundaries and more for data analaysis from R.


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# roadmap
 * import UK MSOA as example and make queries to return boundaries
 * write functions to carryout above
 * generalise even further to import other countries or LSOAs etc

# Dependencies
* mongolite
* geojsonsf

# Import vancouver geojson data
Given a local docker instance of MongoDb

```{bash docker, eval=FALSE}
# on Linux
docker pull mongo
docker run -d -p 27017:27017 --name mongodb mongo
# docker should be running at 27017
```

```{r connect}
library(mongolite)
library(sf)
test = mongo(url = "mongodb://localhost:27017")
class(test)
```

Lets add real data
```{r vancouver}
# geojson file from Uberr
v.path = file.path(tempdir(), "vancouver.json")
if(!exists(v.path)) {
  v.url = "https://github.com/uber-common/deck.gl-data/raw/master/examples/geojson/vancouver-blocks.json"
  download.file(v.url, v.path)
}
v = geojsonsf::geojson_sf(v.path)
nrow(v)
class(v)

# create a collection in your MongoDB instancee
vancouver = mongo("vancouver")
# clear it if has been run before
vancouver$drop()
```

Quietly populate the collection

```{r populatee, results='hide'}

# care is needed to create the mongodb expected geojson objects
by(v, 1:nrow(v), function(x){
  #' geojsonsf package handles Mongo compliant objects in the shape of
  #' {geometry: {type: "Point", coordinates: [0,1]}} and other GeoJSON valid
  #' coordinates and type matching.
  #' unboxed properties
  json = geojsonsf::sf_geojson(x, atomise = TRUE )
  stopifnot(jsonify::validate_json(json))
  vancouver$insert(json)
})
```

Now lets query the collection
```{r queries}
vancouver$find('{}', limit = 1)
vancouver$index((add = '{"geometry" : "2dsphere"}'))

# define a querry as mongolite does not accept somee queries
# a little cheating with copy and paste of one lon lat from the polygons in the data
qry <- '[
  {
    "$geoNear" : { 
      "near" : { "type" : "Point", "coordinates" : [ -123.1107886, 49.2718859 ] },
      "distanceField" : "dist.calculated",
      "maxDistance" : 100,
      "spherical" : true
    }
  }
] '
r = vancouver$aggregate(pipeline = qry)
nrow(r)
# changing the maxDistance will increase the number returned
qry = gsub(x = qry, pattern = "100", replacement = "500")
r = vancouver$aggregate(pipeline = qry)
nrow(r)
```

# public
package no longer private as there is a working function to use.

# Related work
Considered Mapit project, it is great for people who would like to have a web interface, for the use of this package, there is the underlying work of an R interface to MongoDB in `mongolite` and for this use casee, it might be better to use Mongo over PostgreSQL. 

