---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- badges: start -->

# [geocoder](https://layik.github.io/geocoder/) &middot; [![Travis build
status](https://travis-ci.org/layik/geocoder.svg?branch=master)](https://travis-ci.org/layik/geocoder) [![codecov](https://codecov.io/gh/layik/geocoder/branch/master/graph/badge.svg)](https://codecov.io/gh/layik/geocoder) [![Project Status: WIP](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)

Use the power of MongoDB to make spatial queries, find boundaries and more for data analaysis from R.


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# roadmap
 * list of UK geocodes (most used)
 * another country?
 * basic import works with geojson remote/local files, needs  more work
 * find function can query mongodb and return only geometries
 * generalise even further to import other countries or LSOAs etc

# Dependencies
* `mongolite` (main dependency)
* `geojsonsf` (perfect package for link between `sf` and Mongo spatial)
* `jsonify` (helping with validation)
* `sf`

# Import vancouver geojson data
Given a local docker instance of MongoDb

```{bash docker, eval=FALSE}
# on Linux
docker pull mongo
docker run -d -p 27017:27017 --name mongodb mongo
# docker should be running at 27017
```

```{r connect}
library(mongolite)
library(sf)
test = mongo(url = "mongodb://localhost:27017")
class(test)
```

Lets add real data (a small slice of [Uber's Vancouver](https://github.com/uber-common/deck.gl-data/raw/master/examples/geojson/vancouver-blocks.json) dataset.)
```{r vancouver}
# geojson file from Uberr
v.path = file.path(tempdir(), "vancouver.json")
if(!exists(v.path)) {
  v.url = paste0("https://github.com/layik/geocoder/releases/",
                 "download/data/v10.geojson")
  download.file(v.url, v.path)
}
v = geojsonsf::geojson_sf(v.path)
nrow(v)
class(v)

# create a collection in your MongoDB instancee
vancouver = mongo("vancouver")
# clear it if has been run before
vancouver$drop()
vancouver$count()
```

Quietly populate the collection

```{r populatee, results='hide'}
library(geocoder)
gc_import_sf(v, collection = "vancounver")
```

Now lets query the collection
```{r queries}
vancouver$find('{}', limit = 1)
vancouver$index((add = '{"geometry" : "2dsphere"}'))

# define a querry as mongolite does not accept somee queries
# a little cheating with copy and paste of one lon lat from the polygons in the data
qry <- '[
  {
    "$geoNear" : { 
      "near" : { "type" : "Point", "coordinates" : [ -123.1107886, 49.2718859 ] },
      "distanceField" : "dist.calculated",
      "maxDistance" : 3000,
      "spherical" : true
    }
  }
] '
r = vancouver$aggregate(pipeline = qry)
nrow(r)
# changing the maxDistance will increase the number returned
qry = gsub(x = qry, pattern = "3000", replacement = "6000")
r = vancouver$aggregate(pipeline = qry)
nrow(r)
```

# development
To run all the tests, you will require mongo to run. Once you are sure that you have an instance (for example a container running as outliend above), you will need a flag in your R environment `GEOCODER_MONGODB_AVAILABLE=true`. You can use R package `usethis::edit_r_environ()`. 

But the steps should be straightforward for an R developer:

* clone the repo

* from RStudio (cmd/ctrl + T)

* OR from R console:
  + just run `testthat::test_dir("pathToRepo/tests/testthat")`

```{r tests, collapse=TRUE}
library("geocoder")
testthat::test_dir("tests/testthat")
```

# Acknowledgement
This work is funded under the [Turing](https://www.turing.ac.uk/research/research-projects/turing-geovisualization-engine) GeoVisualization Engine project.

# Related work
Considered [Mapit](https://github.com/mysociety/mapit) project, it is great for people who would like to have a web interface, for the use of this package, there is the underlying work of an R interface to MongoDB in `mongolite` and for this use casee, it might be better to use Mongo over PostgreSQL. 

The list of UK geocodes, looks like before the ONS open geography portal, was compiled [here](https://github.com/martinjc/UK-GeoJSON).

