---
title: "Introducing stats19"
author: 
  - "L Hama"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocoder}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
# bibliography: references.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

## Introduction

## UK geocodes case

UK Office of National Statistics Open Geography Portal contains various files which are used by data scientists to plot data on maps. Details of different layers etc.

The MSOA file is [here](https://opendata.arcgis.com/datasets/f341dcfd94284d58aba0a84daf2199e9_0.zip?outSR=%7B%22latestWkid%22%3A27700%2C%22wkid%22%3A27700%7D).
```{r msoa, eval=FALSE}
msoa.folder = file.path(tempdir(), "msoa_folder")
msoa.zip = file.path(tempdir(), "msoa.zip")
if(!exists(msoa.file)) {
  download.file(
    paste0("https://opendata.arcgis.com/datasets/",
           "f341dcfd94284d58aba0a84daf2199e9_0.zip"),
    msoa.zip)
  unzip(msoa.zip, exdir = msoa.folder)
}
# read the shape file using `sf`
msoa.sf = sf::read_sf(file.path(msoa.folder, "Middle_Layer_Super_Output_Areas_December_2001_Full_Clipped_Boundaries_in_England_and_Wales.shp"))
class(msoa.sf)
names(msoa.sf)
substring(geojsonsf::sf_geojson(msoa.sf[1,]), 1, 350)
```

The data is ready in a clean `sf` object and we can convert each row into a ready to be used by `geocder` to import into Mongodb:

```{r eval=FALSE}
geocoder::gc_import_sf(msoa.sf, collection = "msoa")
```

We can now query the database with lists of MSOA codes and easily return GeoJSON formatted "features" ready to be reassembled into `sf`.

```{r, eval=FALSE}
r = gc_find(data.frame(properties.msoa01cd="E02000998"), collection = "msoa")
class(r)
# [1] "data.frame"
substr(r, 1, 200)
# [1] "Feature"
# [2] "list(objectid = 998, msoa01cd = \"E02000998\", msoa01nm = \"Bolton 015\", msoa01nmw = \"Bolton 015\", st_areasha = 1252299.1907256, st_lengths = 8919.80676893236)"
# [3] "list(type = \"MultiPolygon\", coordinates = list(c(-2.45801466425535, -2.4579238289056, -2.45783318819446, -2.45772734343794, -2.4576065864082, -2.45757627561088, -2.45753085804672, -2.45748544050185, -"
```



