---
title: "Introducing geocoder"
author: 
  - "L Hama"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocoder}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

## Introduction
MongoDB comes with built in geospatial indexing and querying [@ameri2014application] around the OGC (Open Geospatial Consortium) standard GeoJSON [@butler2016geojson]. MongoDB accepts GeoJSON formatted coordinates and database indices can be created which fascilitates fast spatial queries supportd by MongoDB. 

A par on SF and R package `sf`.

As a data interchange format, GeoJSON is also widely supported by mapping libraries [@page2015visualising]. In our eAtlas project, we also support the format to visualize and analyse as a native format. GeoJSON of course is JSON which is natively supported by JavaScript and might be the reason for its popularity.

## UK geocodes case

UK Office of National Statistics Open Geography Portal contains various files which are used by data scientists to plot data on maps. Details of different layers etc.

### Round trip through MongoDB
To showcasee the capabilities of `geocoder` package, we can do a round trip data processing from a seudo-standard R spatial `sf` class back to itself through MongoDB's geospatial querying. That is:

* take a shapefile from UK's Open Geography Portal

* turn it into an `sf` object

* write it into a MongoDB collection and create a spatial index on it.

* reassemble the output from find queries back to an `sf` object.

One of the UK census boundaries is called Middle Superr Output Area (MSOA), 
which can be found 
[here](https://opendata.arcgis.com/datasets/f341dcfd94284d58aba0a84daf2199e9_0.zip).

```{r msoa, eval=FALSE}
msoa.folder = file.path(tempdir(), "msoa_folder")
msoa.zip = file.path(tempdir(), "msoa.zip")
if(!exists(msoa.file)) {
  download.file(
    paste0("https://opendata.arcgis.com/datasets/",
           "f341dcfd94284d58aba0a84daf2199e9_0.zip"),
    msoa.zip)
  unzip(msoa.zip, exdir = msoa.folder)
}
# read the shape file using `sf`
msoa.sf = sf::read_sf(
  list.files(msoa.folder, pattern = "England_and_Wales.shp"))
class(msoa.sf)
# substring(geojsonsf::sf_geojson(msoa.sf[1,]), 1, 350)
```

The data is ready in a clean `sf` object and we can convert each row into a ready to be used by `geocder` to import into Mongodb:

```{r eval=FALSE}
geocoder::gc_import_sf(msoa.sf, collection = "msoa")
```

We can now query the database with lists of MSOA codes and easily return GeoJSON formatted "features" ready to be reassembled into `sf`.

```{r, eval=FALSE}
r = gc_find(data.frame(properties.msoa01cd="E02000998"), collection = "msoa")
class(r)
# [1] "data.frame"
substr(r, 1, 200)
# [1] "Feature"
# [2] "list(objectid = 998, msoa01cd = \"E02000998\", msoa01nm = \"Bolton 015\", msoa01nmw = \"Bolton 015\", st_areasha = 1252299.1907256, st_lengths = 8919.80676893236)"
# [3] "list(type = \"MultiPolygon\", coordinates = list(c(-2.45801466425535, -2.4579238289056, -2.45783318819446, -2.45772734343794, -2.4576065864082, -2.45757627561088, -2.45753085804672, -2.45748544050185, -"

```

A reproducible (as the UK msoa example is too large) for this Rmarkdown document would be a slice of Uber's Vancouver land area price.

```{r reproducible}
# geojson file from Uber
library(sf)
v.path = file.path(tempdir(), "vancouver.json")
if(!exists(v.path)) {
  v.url = paste0("https://github.com/layik/geocoder/releases/",
                 "download/data/v10.geojson")
  download.file(v.url, v.path)
}
v = geojsonsf::geojson_sf(v.path)
st_drop_geometry(v[8,])
vancouver = mongolite::mongo("vancouver")
plot(v[8, 2])
```

Let us retrieve the data and plot the land:
```{r reproducible2}
vancouver$drop() # if exists
geocoder::gc_import_sf(v, collection = "vancouver")
r = vancouver$iterate(query = '{"properties.growth":0.2781}')
r.sf = geojsonsf::geojson_sf(r$json())
plot(r.sf[, 2])
```


## References


